---
title: "Super Bowl Brands' Videos Engagement Analysis"
author: "Huwail Alantari, Vishwa Banker, Emily Boyer, Chloe Li"
date: "April 9, 2016"
output: 
  ioslides_presentation:
    transition: faster
    smaller: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Intro
- Quantify amount of engagement and views before (1/11-2/6) and after (2/6-2/13) the Super Bowl
- Calculated amount of engagement increased/decreased during each period (total of 5 periods)
- Sum the amount of engagement by type of the engagement, by brand, by official channel
- Plots trend (time series) through all periods

```{r, error=FALSE, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls())

#load library
library(dplyr)
library(ggplot2)


setwd('/Users/ChloeLi/Documents/MS Business Analytics/Digital Marketing/Project Due April14/MBAX-6350-MS-ZEFR-project-data-2016')

#load csv.  file per brand
#these files contains amount of engagement increased/decreased per period, and separated by official brand
MountainDew <- read.csv('MountainDew_perid.csv')
#Reese <- read.csv('Reese_period.csv')
#ShockTop <- read.csv('ShockTop_period.csv')
#Snickers <- read.csv('Snickers_period.csv')


#-------------------------------------------plot mountain dew----------------------------------------------#
#grap all engagement variables for calculating the sum
#this is total engagemet per period WITHOUT identifing official channel
mydata_total <- select(MountainDew, -X,-video_id, -channel_id,-Brand, -officialBrandContent)



#Comments_official <- select(Official, Comments1_office = period1Comments, Comments2_office=period2Comments, Comments3_office=period3Comments, 
#                   Comments4_office=period4Comments,Comments5_office=period5Comments)


#sum of engagement period for all channels
m <- as.data.frame(apply(mydata_total, 2, FUN = sum))
#rename column name
colnames(m) <- "Total_Engagement"
m$period <- c("1/11-1/16","1/16-1/23","1/23-1/30","1/30-2/6","2/6-2/13")
m$Engagement <- c("period1Comments", "period2Comments", "period3Comments",
 "period4Comments", "period5Comments", "period1Likes",   
 "period2Likes",    "period3Likes",    "period4Likes" , "period5Likes",    "period1Disikes",  "period2Dislikes","period3Dislikes", "period4Dislikes", "period5Dislikes", "period1Views",    "period2Views",    "period3Views","period4Views", "period5Views")

#drop row names
row.names(m) = NULL
#rearrange
m <- select(m, Period = period, Engagement, Total_Engagement)

#create official channel engagement
#grap all engagement variables also grap all official channel
Official <- filter(MountainDew, officialBrandContent == TRUE)
Official <- select(Official, -X,-video_id, -channel_id,-Brand, -officialBrandContent)

n <- as.data.frame(apply(Official, 2, FUN = sum))
colnames(n) <- "Official_Engagement"
m$Official_Engagement <- n$Official_Engagement
#calculate engagement for unofficial channel
m$unOfficial_Engagement <- m$Total_Engagement - m$Official_Engagement

library(reshape2)
#transform wide to long format
Long_m <- melt(m, id.vars = c("Period", "Engagement"))
Long_m <- select(Long_m, Period,Engagement_Type = Engagement, Channel_Type = variable, Engagement_Num = value)
```
## Trend for Number of Comments By Channel Type - Mountain Dew

```{r, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
#-------------------------------------------plot mountain dew----------------------------------------------#
#for comments

Comments_MD <- Long_m[grep("Comments", Long_m$Engagement_Type), ]

ggplot(data=Comments_MD, aes(x=Period, y=Engagement_Num, group = Channel_Type, colour = Channel_Type)) +
    geom_line(size=0.75) +
    geom_vline(xintercept = 4.5, linetype = "longdash") +
    xlab("Number of Comments") +
    ggtitle("Comments by Channel Type for Mountains Dew")

```

## Trend for Number of Likes By Channel Type - Mountain Dew

```{r, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
#-------------------------------------------plot mountain dew----------------------------------------------#
#for comments

Likes_MD <- Long_m[grep("Likes", Long_m$Engagement_Type), ]

ggplot(data=Likes_MD, aes(x=Period, y=Engagement_Num, group = Channel_Type, colour = Channel_Type)) +
    geom_line(size=0.75) +
    geom_vline(xintercept = 4.5, linetype = "longdash") +
    xlab("Number of Likes") +
    ggtitle("Likes by Channel Type for Mountains Dew")

```

## Trend for Number of Dislikes By Channel Type - Mountain Dew

```{r, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
#-------------------------------------------plot mountain dew----------------------------------------------#
#for comments

Dislikes_MD <- Long_m[grep("Disikes|Dislikes", Long_m$Engagement_Type), ]

ggplot(data=Dislikes_MD, aes(x=Period, y=Engagement_Num, group = Channel_Type, colour = Channel_Type)) +
    geom_line(size=0.75) +
    geom_vline(xintercept = 4.5, linetype = "longdash") +
    xlab("Number of Dislikes") +
    ggtitle("Dislikes by Channel Type for Mountains Dew")

```

## Trend for Number of Views/Exposure By Channel Type - Mountain Dew

```{r, echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}
#-------------------------------------------plot mountain dew----------------------------------------------#
#for comments

Views_MD <- Long_m[grep("Views", Long_m$Engagement_Type), ]

ggplot(data=Views_MD, aes(x=Period, y=Engagement_Num, group = Channel_Type, colour = Channel_Type)) +
    geom_line(size=0.75) +
    geom_vline(xintercept = 4.5, linetype = "longdash") +
    xlab("Number of Exposure") +
    ggtitle("Exposure by Channel Type for Mountains Dew")

```






